### 2018-08-24 doc/large_strain.tex
add more caching algorithms plus minor cleanup

% \usepackage[linesnumbered,lined,commentsnumbered,ruled]{algorithm2e}
%\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{bm}
%\usepackage{bbm}
%\usepackage{caption}
%\usepackage{color}
%\usepackage{geometry}
%\usepackage{graphicx}
%\usepackage[colorlinks,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
%\usepackage{layouts}
%\usepackage{mathrsfs}
%\usepackage{pgfkeys,pgfmath,pgfcore}
%\usepackage{srcltx}
%\usepackage[labelformat=simple]{subcaption}
%\usepackage{xspace}

\usepackage[linesnumbered,lined,commentsnumbered,ruled]{algorithm2e}

@@

@@

\begin{algorithm}[h]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$, current FE solution $\gz u$,
  cacheed $c_1 := \mu - 2 \lambda \log(J)$ for each cell and quadrature point}
  \Output{action of the FE tangent on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $\gz F = \gz I + \gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $J = \rm{det}(\gz F)$ \tcp*{scalar}
          evaluate $\gz b = \gz F \cdot \gz F^T$ \tcp*{2nd order symmetric}
          evaluate $\gz x_1 := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\gz \tau = \mu \gz b - c_1 \gz I$ \tcp*{2nd order symmetric}
          evaluate $J \boldsymbol{\mathcal{G}} = x_1 \cdot \gz \tau$ \tcp*{2nd order symmetric}
          ``submit'' contraction $\left[\gz \nabla \gz N_i : J\boldsymbol{\mathcal{G}}\right] / J$\;
          evaluate $\gz x_2 := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $J \boldsymbol{\mathcal{C}} = 2 c_1 \gz x_2 + 2 \lambda \rm{tr}( \gz x_2) \gz I $ \tcp*{2nd order symmetric}
          ``submit'' contraction $\left[\gz \nabla_{sym} \gz N_i : J \boldsymbol{\mathcal{C}} \right]/ J$ \;
        }
  }
  \caption{Matrix-free tangent cell operator, variant 1: minimum caching.}
  \label{alg:mf_multivector}
\end{algorithm}

>>>>

% \usepackage[linesnumbered,lined,commentsnumbered,ruled]{algorithm2e}
%\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{bm}
%\usepackage{bbm}
%\usepackage{caption}
\usepackage{color}
%\usepackage{geometry}
%\usepackage{graphicx}
%\usepackage[colorlinks,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
%\usepackage{layouts}
%\usepackage{mathrsfs}
%\usepackage{pgfkeys,pgfmath,pgfcore}
%\usepackage{srcltx}
%\usepackage[labelformat=simple]{subcaption}
%\usepackage{xspace}
\usepackage[linesnumbered,lined,commentsnumbered,ruled]{algorithm2e}

@@

{\color{red}
[DA] Daniel, please add a brief intro about MF (readership will likely not know anything about it), maybe borrow 1-2 pictures to explain how tensor-product structure is handled.
}

@@

\begin{algorithm}[h]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$, current FE solution $\gz u$,
  cacheed $c_1 := \mu - 2 \lambda \log(J)$ for each cell and quadrature point}
  \Output{action of the FE tangent on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $\gz F = \gz I + \gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $J = \rm{det}(\gz F)$ \tcp*{scalar}
          evaluate $\gz b = \gz F \cdot \gz F^T$ \tcp*{2nd order symmetric}
          evaluate $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\gz \tau = \mu \gz b - c_1 \gz I$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{G}} \gz g = \gz g \cdot \gz \tau/J$ \tcp*{2nd order}
          ``submit'' contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$\;
          evaluate $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{C}}\gz g_s = \left[2 c_1 \gz g_s + 2 \lambda\, \rm{tr}( \gz g_s) \gz I \right]/J$ \tcp*{2nd order symmetric}
          ``submit'' contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$ \;
        }
  }
  \caption{Matrix-free tangent cell operator: scalar cache}
  \label{alg:mf_scalar}
\end{algorithm}

\begin{algorithm}[h]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached
  $c_1 := 2\left[\mu - 2 \lambda \log(J)\right]/J$,
  $c_2 := 2\lambda/J$ and $\gz \tau/J$ for each cell and quadrature point}
  \Output{action of the FE tangent on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\boldsymbol{\mathcal{G}}\gz g = \gz g \cdot \left[\gz \tau/J \right]$ \tcp*{2nd order}
          ``submit'' contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$ \;
          evaluate $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{C}} \gz g_s = c_1 \gz g_s + c_2 \, \rm{tr}(\gz g_s) \gz I$ \tcp*{2nd order symmetric}
          ``submit'' contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}} \gz g_s $ \;
        }
  }
  \caption{Matrix-free tangent cell operator, variant 1: cache second order Kirchhoff stress $\gz \tau$ and thereby avoid the need to evaluate refential quantities like $\gz F$ at runtime.}
  \label{alg:mf_tensor2}
\end{algorithm}

\begin{algorithm}[h]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached $\gz \tau/J$ and $\boldsymbol{\mathcal{C}}$ for each cell and quadrature point}
  \Output{action of the FE tangent on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\boldsymbol{\mathcal{G}} \gz g = \gz g \cdot \left[\gz \tau/J\right]$ \tcp*{2nd order}
          ``submit'' contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$\;
          evaluate $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{C}}\gz g_s = \boldsymbol{\mathcal{C}}\gz : \gz g_s$ \tcp*{2nd order symmetric}
          ``submit'' contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$ \;
        }
  }
  \caption{Matrix-free tangent cell operator: cache fourth order material tangent $\boldsymbol{\mathcal{C}}$ and Kirchhoff stress $\gz \tau$.}
  \label{alg:mf_tensor4}
\end{algorithm}

The primary difference between Algorithm \ref{alg:mf_tensor2} and Algorithm \ref{alg:mf_tensor4}, is that the former utilizes the chosen structure of free energy function and thereby ends up with a computationally cheap way of evaluation of $\boldsymbol{\mathcal{C}} \gz g_s $, whereas the latter performs the double contraction between the fourth order and second order tensors.
